---
title: "Experiment 2 - Rosenbaum, Grassie & Hartley"
author: "Gail Rosenbaum"
date: "01/15/2022"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    df_print: kable
---

<style type="text/css">

h1.title {
 font-size: 38px;
}
h1 { /* Header 1 */
 font-size: 28px;
}
h2 { /* Header 2 */
   font-size: 22px;
}
h3 { /* Header 3 */
 font-size: 18px;
}

</style>

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, cache = FALSE, 
                      message = FALSE, results = "asis", 
                      warning = FALSE, include = TRUE)

```

# Setup
## Load Libraries
```{r libraries, include=FALSE}

library(knitr)
library(tidyverse)
library(pander)
library(lme4)
library(sjPlot)
library(here)

```

Note: Experiment 2 data were originally published in the following paper:
Rouhani, N., Norman, K. A., & Niv, Y. (2018). Dissociable effects of surprising rewards on learning and memory. Journal of Experimental Psychology: Learning, Memory, and Cognition, 44(9), 1430–1443. https://doi.org/10.1037/xlm0000518


## Load data
```{r loadData, include=FALSE}

# load trial-by-trial memory data
MemDF_Full <- read_csv(here("data_Exp2", "memory_allExps"))

#load memory summary statistics, including false alarms so we can include that in models
FA_Full <- read_csv(here("data_Exp2", "indDiff_allExps"))

#load a dataframe with best-fit RL model parameters, which we'll use to compute asymmetry index (AI)
AIDF_Full <- read_csv(here("data_Exp2", "BestFitModelParameters11272020.csv"))

# Load trial-by-trial RPE estimates from the models
RPEDF_Full <- read_csv(here("data_Exp2", "TrialByTrialRPEs_11272020.csv"))

```

## Set theme for sjPlot
```{r setSJPlotTheme, include=FALSE}

set_theme(
  base = theme_bw(), 
  theme.font = "sans",
  panel.major.gridcol = "white",
  panel.minor.gridcol = "white", 
  legend.backgroundcol = "white",
  legend.item.backcol = "white",
  legend.item.size = 1
)

```

## Setup
```{r computeMerge, include=FALSE}

#Scale FA rate
FA_Full$FAScale <- scale(FA_Full$FARate)

#Merge RPE data with trial-by-trial memory data with h
MemDF2_Full <- merge(RPEDF_Full, MemDF_Full, by.x = c("subj","trialExp"),
                     by.y = c("subj","trialLearnExp"))

#merge memory summary data with the larger dataset
MemDF2_Full <- merge(MemDF2_Full, FA_Full, by = "subj")

#set up variables for analysis
MemDF2_Full$PositiveRPE <- factor(MemDF2_Full$PositiveRPE)

MemDF2_Full$AIScale <- scale(MemDF2_Full$AsymmIdx)

MemDF2_Full$TrialScale <- scale(MemDF2_Full$trialMemory)

MemDF2_Full$AbsRPEScale <- scale(MemDF2_Full$AbsRPE)

MemDF2_Full$PositiveRPEC <- ifelse(MemDF2_Full$PositiveRPE == 1, 1,
                                   ifelse(MemDF2_Full$PositiveRPE == 0, -1, NA))

#merge FAs with asymmetry index
AIDF_Full <- merge(AIDF_Full, FA_Full[,c("subj","exp")],by = "subj")

```

# Assess model fit
```{r Correlations}

#for each subject, find the correlation between the modeled image estimates and the subjects' actual estimates
corrsFull <- RPEDF_Full %>%
    group_by(subj) %>%
    summarise(correlationFull =  cor(estimate, modval))

#merge correlation variables with other datasets
corrsFull <- merge(corrsFull, FA_Full[,c("subj","exp")], by = "subj")
corrsFull <- merge(corrsFull, AIDF_Full, by = c("subj","exp"))

```


In the full dataset, the correlation between the model’s value estimates and 
actual value estimates from participants’ task experience: 
mean = `r round(mean(corrsFull$correlationFull),2)`
SD = `r round(sd(corrsFull$correlationFull),2)`

# Compute asymmetry index 

```{r AI}

#compute AI 
AIDF_Full$AI <- (AIDF_Full$alphapos - AIDF_Full$alphaneg)/
    (AIDF_Full$alphapos + AIDF_Full$alphaneg) 

```

Asymmetry index:
Mean = `r round(mean(AIDF_Full$AI),2)`
SD = `r round(sd(AIDF_Full$AI),2)`


# Identify participants with good vs. subpar model fit

```{r subsetGoodFits}

#find the subs who have a correlation greater than 0 in all models, and with 
#nloglik less than 0
corrsFull$GoodFit <- ifelse(corrsFull$correlationFull > 0 & 
                                corrsFull$nloglik < 0, TRUE, FALSE)

corrsFull_GF <- subset(corrsFull, corrsFull$GoodFit ==TRUE)

```


Number of participants total (regardless of fit) = `r nrow(corrsFull)`


Number of participants with correlation lower that 0 or negative loglikelihood 
less than 0 (i.e., poorly fit by the model) = 
`r sum(corrsFull$GoodFit == FALSE) `


Number of participants with adequate model fit = `r nrow(corrsFull_GF)`


# Analyses with participants with good model fit (main text analysis)

In the subset of data with good RL model fit, the correlation between the model’s 
value estimates and actual value estimates from participants’ task experience: 
mean = `r round(mean(corrsFull_GF$correlationFull),2)`
SD = `r round(sd(corrsFull_GF$correlationFull),2)`


## Clean up "good fit" dataset

```{r cleanupgoodfit}

#get list of subs with a good fit
GoodFit <- subset(corrsFull$subj, corrsFull$GoodFit == TRUE)

#subset AI dataset
AIDF_Full_GF <- subset(AIDF_Full, AIDF_Full$subj %in% GoodFit)

#subset Mem dataset
MemDF2_Full_GF <- subset(MemDF2_Full, MemDF2_Full$subj %in% GoodFit)

#rescaling variables because we removed some subs
MemDF2_Full_GF$AIScale <- scale(MemDF2_Full_GF$AsymmIdx)

MemDF2_Full_GF$TrialScale <- scale(MemDF2_Full_GF$trialMemory)

MemDF2_Full_GF$AbsRPEScale <- scale(MemDF2_Full_GF$AbsRPE)

MemDF2_Full_GF$FAScale <- scale(MemDF2_Full_GF$FARate)
```


## Plot parameter distributions


Appendix 1—figure 14A-C

```{r goodFitparams}

ggplot(AIDF_Full_GF, aes(x=alphapos)) + 
    geom_histogram(fill = c("#6baed6"))+ 
    xlab(expression(paste(alpha,"+"))) +
    theme_bw() + 
    theme(text=element_text(family="sans",size=16),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())

ggplot(AIDF_Full_GF, aes(x=alphaneg)) + 
    geom_histogram(fill = c("#6baed6")) + 
    xlab(expression(paste(alpha,"-"))) +
    theme_bw() + 
    theme(text=element_text(family="sans",size=16),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())

ggplot(AIDF_Full_GF, aes(x=AI)) + 
    geom_histogram(fill = c("#6baed6")) +
    xlab("Asymmetry Index (AI)")+
    theme_bw() + 
    theme(text=element_text(family="sans",size=16),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())

```


Asymmetry index in subjects with good RL model fit:
Mean = `r round(mean(AIDF_Full_GF$AI),2)`
SD = `r round(sd(AIDF_Full_GF$AI),2)`



## Memory Mixed-Effects Regression
### Model

```{r glmergoodfit}

#run glmer model with good-fit subjects

lmAIRPEMaxGF <- glmer(
    hit ~ AIScale*AbsRPEScale*PositiveRPEC+TrialScale+FAScale+
        (1+AbsRPEScale +PositiveRPEC+TrialScale || subj), 
    family=binomial, data=MemDF2_Full_GF,
    control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1e6)))

tab_model(lmAIRPEMaxGF, show.stat = TRUE, 
                           show.df = TRUE, string.stat = "z",
                           col.order = c("stat","p","est", "ci")) 
```


### Plot


Figure 6A

```{r fig6a}

#Plot odds ratios
plotlabels <- c("Asymmetry Index (AI)", "PE Magnitude","PE Valence",
                "Memory Trial Number","False Alarm Rate",
                "AI:PE Magnitude","AI:PE Valence","PE Magnitude:PE Valence",
                "AI:PEMagnitude:PEValence")
names(plotlabels) <- c("AIScale", "AbsRPEScale","PositiveRPEC",
                       "TrialScale","FAScale",
                       "AIScale:AbsRPEScale","AIScale:PositiveRPEC",
                       "AbsRPEScale:PositiveRPEC",
                       "AIScale:AbsRPEScale:PositiveRPEC")
threewayintxplotPaper <- plot_model(lmAIRPEMaxGF, colors = "bw", 
                                    show.values = TRUE, 
                                    value.offset = .4,  
                                    order.terms=c(4,5,1,2,3,6,7,8,9),
                                    title = "Fixed Effects",
                                    axis.labels = plotlabels, 
                                    vline.color = "grey")
threewayintxplotPaper + 
    theme(text=element_text(family="sans",size=12),
          axis.title=element_text(size=16),
          panel.background = element_rect(fill = "white",  colour = "white"), 
          panel.border = element_rect(colour = "black", fill=NA)) +
    scale_y_log10(limits = c(.5,2))


```


Figure 6B

```{r fig6b}

#running a new model with non-centered variables to easily plot the 3-way intx
#didn't have a factor valence variable in this df so creating one
MemDF2_Full_GF$PositiveRPEFact <- factor(MemDF2_Full_GF$PositiveRPE, 
                                         levels = c(0,1), 
                                         labels = c("Negative Prediction Error",
                                                    "Positive Prediction Error"))

#run new model 
lmAIRPEMaxGFPlot <- glmer(
    hit ~ AsymmIdx * AbsRPE * PositiveRPEFact + TrialScale + FARate + 
        (1 + AbsRPEScale + PositiveRPEC + TrialScale || subj), 
    family=binomial, data=MemDF2_Full_GF,
    control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1e6)))

#make 3-way interaction plot
AIRPEFig <- plot_model(lmAIRPEMaxGFPlot,
           type = "pred", 
           terms = c("AbsRPE [all]", "AsymmIdx [-.8,0,.8]", "PositiveRPEFact"),
           colors = c("#bdd7e7","#6baed6","#2171b5")) +
    scale_y_continuous(breaks=c(.4,.5,.6,.7,.8,.9), 
                       labels = c("40%","50%","60%","70%","80%","90%"))
AIRPEFig <- AIRPEFig + theme(
        text=element_text(family="sans",size=12),
        strip.background =element_rect(fill="white"),
        panel.spacing = unit(1, "lines"))
        
AIRPEFig <- AIRPEFig + labs(title="Predicted Memory Accuracy by Choice Context",
        x="Absolute Value of Prediction Error",
        y= "Estimated Marginal Means for \n% Correct Memory Responses",
        linetype="Asymmetry\nIndex",
        color = "Asymmetry\nIndex")
AIRPEFig


```


Appendix 1—-figure 13D

```{r figa13D}

#reformat Figure 6B a bit for the appendix figure
AIRPEFigAppendix <- AIRPEFig + labs(title="Experiment 2 - Explicit Predictions",
        x="Absolute Value of Prediction Error",
        y= "Estimated Marginal Means for \n% Correct Memory Responses",
        linetype="Asymmetry\nIndex",
        color = "Asymmetry\nIndex") + 
    theme_bw() + 
    theme(
        text=element_text(family="sans",size=12),
        panel.background = element_rect(fill = "white",  colour = "white"),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.spacing = unit(1, "lines"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background =element_rect(fill="white"), 
        legend.position = "none")
AIRPEFigAppendix 

```


# Sensitivity analysis (Appendix 1)

Rerunning the mixed-effects model with the full dataset, including those with poor model fit.


## Model

```{r sensitivityanalysis}

#glmer model including all participants, regardless of model fit
#the highest level model that converged (maximal model did not converge)
lmAIRPEMax <- glmer(
    hit ~ AIScale*AbsRPEScale*PositiveRPEC+TrialScale+FAScale+ 
        (1+AbsRPEScale +TrialScale || subj), 
    family=binomial, data=MemDF2_Full,
    control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1e6)))

#table (for some reason using tab_model works in this script but not in the Exp1 
#script; In that script I have to use the glmerReport function I made)
tab_model(lmAIRPEMax, show.stat = TRUE, 
                           show.df = TRUE, string.stat = "z",
                           col.order = c("stat","p","est", "ci"))

```


## Plot


Appendix 1—figure 15A

```{r figa15a}

#Plot odds ratios
plotlabelsall <- c("Asymmetry Index (AI)", 
                   "PE Valence",
                   "PE Magnitude",
                   "Memory Trial Number",
                   "False Alarm Rate", 
                   "AI:PE Magnitude",
                   "AI:PE Valence",
                   "PE Magnitude:PE Valence",
                   "AI:PE Magnitude:PE Valence")
names(plotlabelsall) <- c("AIScale",
                        "PositiveRPEC",
                        "AbsRPEScale",
                        "TrialScale",
                        "FAScale",
                        "AIScale:PositiveRPEC",
                        "AIScale:AbsRPEScale",
                        "PositiveRPEC:AbsRPEScale",
                        "AIScale:PositiveRPEC:AbsRPEScale")
threewayintxplotall <- plot_model(
    lmAIRPEMax, colors = "bw", show.values = TRUE, 
    value.offset = .4,  order.terms=c(4,5,1,2,3,6,7,8,9),
    axis.labels = plotlabelsall, title = "Fixed Effects", 
                                    vline.color = "grey") 
threewayintxplotall + 
    ylab("Odds of Correct Memory Response") + 
    theme(text=element_text(family="sans",size=12),
          axis.title=element_text(size=16),
          panel.background = element_rect(fill = "white",  colour = "white"), 
          panel.border = element_rect(colour = "black", fill=NA)) +
    scale_y_log10(limits = c(.5,2))


    
    
```


Appendix 1—figure 15B

```{r figa15b}

#make a factor instead of contrast for easier interaction plotting
MemDF2_Full$PositiveRPEFact <- factor(
    MemDF2_Full$PositiveRPE, 
    levels = c(0,1), 
    labels = c("Negative Prediction Error","Positive Prediction Error"))

#model with unscaled variables for intx plot
lmAIRPEMaxPlot <- glmer(
    hit ~ AsymmIdx*PositiveRPEFact*AbsRPE+TrialScale+ FAScale+
        (1+AbsRPE+TrialScale || subj), 
    family=binomial, data=MemDF2_Full,
    control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1e6)))

AIRPEFig <- plot_model(lmAIRPEMaxPlot,
           type = "pred", 
           terms = c("AbsRPE [all]", "AsymmIdx [-.8,0,.8]", "PositiveRPEFact"),
           colors = c("#bdd7e7","#6baed6","#2171b5")) +
    scale_y_continuous(breaks=c(.4,.5,.6,.7,.8,.9), 
                       labels = c("40%","50%","60%","70%","80%","90%"))

AIRPEFig <- AIRPEFig + theme(
        text=element_text(family="sans",size=12),
        strip.background =element_rect(fill="white"),
        panel.spacing = unit(1, "lines"))
        
AIRPEFig <- AIRPEFig + labs(title="Predicted Memory Accuracy by Choice Context",
        x="Absolute Value of Prediction Error",
        y= "Estimated Marginal Means for \n% Correct Memory Responses",
        linetype="Asymmetry\nIndex",
        color = "Asymmetry\nIndex")
AIRPEFig+ 
    theme_bw() + 
    theme(
        text=element_text(family="sans",size=12),
        panel.background = element_rect(fill = "white",  colour = "white"),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.spacing = unit(1, "lines"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_rect(fill="white"))



```


Appendix 1—figure 15C

```{r figa15c}

#2-way interaction plot (the significant effect in this model)
plot_model(lmAIRPEMaxPlot,type = "pred",
           terms = c("PositiveRPEFact", "AsymmIdx [-.8,0,.8]"), 
           colors = c("#bdd7e7","#6baed6","#2171b5"))+ 
    labs(title="Predicted Memory Accuracy by Choice Context",
        x="Absolute Value of Prediction Error",
        y= "Estimated Marginal Means for \n% Correct Memory Responses",
        linetype="Asymmetry\nIndex",
        color = "Asymmetry\nIndex")+ 
    theme_bw() + 
    theme(
        text=element_text(family="sans",size=12),
        panel.background = element_rect(fill = "white",  colour = "white"),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.spacing = unit(1, "lines"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        strip.background = element_rect(fill="white"))

```

