---
title: "Experiment 2 - Rosenbaum, Grassie & Hartley"
author: "Gail Rosenbaum"
date: "11/27/2020"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    df_print: kable
---

<style type="text/css">

h1.title {
 font-size: 38px;
}
h1 { /* Header 1 */
 font-size: 28px;
}
h2 { /* Header 2 */
   font-size: 22px;
}
h3 { /* Header 3 */
 font-size: 18px;
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
## Load Libraries
```{r libraries, include=FALSE}

library(knitr)
library(tidyverse)
library(pander)
library(lme4)
library(lmerTest)
library(sjPlot)
library(dplyr)
library(car)
library(gridExtra)
library(reshape2)
library(ggpubr)
library(cowplot)
library(ggeffects)
library(ggthemes)
library(RColorBrewer)
library(here)
library(GGally)
library(emmeans)
library(sjmisc)
library(sjstats)
library(r2glmm)
library(effectsize)
library(grid)
library(ggforce)
library(broom)

```

Note: Experiment 2 data were originally published in the following paper:
Rouhani, N., Norman, K. A., & Niv, Y. (2018). Dissociable effects of surprising rewards on learning and memory. Journal of Experimental Psychology: Learning, Memory, and Cognition, 44(9), 1430–1443. https://doi.org/10.1037/xlm0000518


## load data
```{r loadData, include=FALSE}

# load memory data - Regression version (based on Jian's script)
MemDF_Full <- read_csv(here("data_Exp2", "memory_allExps"))
FA_Full <- read_csv(here("data_Exp2", "indDiff_allExps"))

AIDF_Full <- read_csv(here("data_Exp2", "BestFitModelParameters11272020.csv"))

# Load trial-by-trial RPE estimates from the models
RPEDF_Full <- read_csv(here("data_Exp2", "TrialByTrialRPEs_11272020.csv"))


```

## Set theme for sjPlot
```{r setSJPlotTheme, include=FALSE}
# set_theme(
#   base = theme_bw(), 
#   theme.font = "sans",
#   panel.backcol = "white",
#   panel.gridcol = "white",
#   panel.col = "white",
#   panel.bordercol = "black",
#   plot.backcol = "white",
#   legend.backgroundcol = "white",
#   legend.item.backcol = "white",
#   legend.item.size = 1
# )

set_theme(
  base = theme_bw(), 
  theme.font = "sans",
  panel.major.gridcol = "white",
  panel.minor.gridcol = "white", 
  legend.backgroundcol = "white",
  legend.item.backcol = "white",
  legend.item.size = 1
)

```


# Compute new variables and merge
```{r computeMerge, include=FALSE}

FA_Full$FAScale <- scale(FA_Full$FARate)

MemDF2_Full <- merge(RPEDF_Full, MemDF_Full, by.x = c("subj","trialExp"),by.y = c("subj","trialLearnExp"))

MemDF2_Full <- merge(MemDF2_Full, FA_Full, by = "subj")

MemDF2_Full$PositiveRPE <- factor(MemDF2_Full$PositiveRPE)

MemDF2_Full$AIScale <- scale(MemDF2_Full$AsymmIdx)

MemDF2_Full$TrialScale <- scale(MemDF2_Full$trialMemory)

MemDF2_Full$AbsRPEScale <- scale(MemDF2_Full$AbsRPE)

MemDF2_Full$PositiveRPEC <- ifelse(MemDF2_Full$PositiveRPE == 1, 1, ifelse(MemDF2_Full$PositiveRPE == 0, -1, NA))

AIDF_Full <- merge(AIDF_Full, FA_Full[,c("subj","exp")],by = "subj")

```

# Correlations between subject and model estimates
```{r Correlations}

#for each subject, find the correlation between the modeled image estimates and the subjects' actual estimates
corrsFull <- RPEDF_Full %>%
    group_by(subj) %>%
    summarise(correlationFull =  cor(estimate, modval))

#merge correlation variables with other datasets
corrsFull <- merge(corrsFull, FA_Full[,c("subj","exp")], by = "subj")
corrsFull <- merge(corrsFull, AIDF_Full, by = c("subj","exp"))

#mean/SD of correlation with all participants in the sample
mean(corrsFull$correlationFull)
sd(corrsFull$correlationFull)

```

# Compute asymmetry index 
```{r AI}

#compute AI
AIDF_Full$AI <- (AIDF_Full$alphapos - AIDF_Full$alphaneg)/(AIDF_Full$alphapos + AIDF_Full$alphaneg) 

#summary stats
mean(AIDF_Full$AI)
sd(AIDF_Full$AI)


```

# glmer
```{r glmerModelEstimates}

#the model that converges
lmAIRPEMax <- glmer(hit ~ AIScale*AbsRPEScale*PositiveRPEC+TrialScale+FAScale+ (1+AbsRPEScale +TrialScale || subj), family=binomial, data=MemDF2_Full,control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1e6)))

#Plot odds ratios
plotlabelsall <- c("Asymmetry Index (AI)", 
                   "PE Valence",
                   "PE Magnitude",
                   "Memory Trial Number",
                   "False Alarm Rate", 
                   "AI:PE Magnitude",
                   "AI:PE Valence",
                   "PE Magnitude:PE Valence",
                   "AI:PE Magnitude:PE Valence")
names(plotlabelsall) <- c("AIScale",
                        "PositiveRPEC",
                        "AbsRPEScale",
                        "TrialScale",
                        "FAScale",
                        "AIScale:PositiveRPEC",
                        "AIScale:AbsRPEScale",
                        "PositiveRPEC:AbsRPEScale",
                        "AIScale:PositiveRPEC:AbsRPEScale")
threewayintxplotall <- plot_model(lmAIRPEMax, colors = "bw", show.values = TRUE, 
                                    value.offset = .4,  order.terms=c(4,5,1,2,3,6,7,8,9),
                                    axis.labels = plotlabelsall, title = "Fixed Effects", 
                                    vline.color = "grey")
threewayintxplotall + ylab("Odds of Correct Memory Response")+ theme(
        text=element_text(family="sans",size=12))  + scale_y_log10(limits = c(.5,2))

#table (for some reason using tab_model works in this script but not in the Exp1 script; In that script I have to use the glmerReport function I made)
tab_model(lmAIRPEMax, show.stat = TRUE, 
                           show.df = TRUE, string.stat = "z",
                           col.order = c("stat","p","est", "ci"))

#make a factor instead of contrast for easier interaction plotting
MemDF2_Full$PositiveRPEFact <- factor(MemDF2_Full$PositiveRPE, levels = c(0,1), labels = c("Negative Prediction Error","Positive Prediction Error"))

#model with unscaled variables for intx plot
lmAIRPEMaxPlot <- glmer(hit ~ AsymmIdx*PositiveRPEFact*AbsRPE+TrialScale+ FAScale+ (1+AbsRPE+TrialScale || subj), family=binomial, data=MemDF2_Full,control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1e6)))

AIRPEFig <- plot_model(lmAIRPEMaxPlot,
           type = "pred", 
           terms = c("AbsRPE [all]", "AsymmIdx [-.8,0,.8]", "PositiveRPEFact"),
           colors = c("#bdd7e7","#6baed6","#2171b5")) +
    scale_y_continuous(breaks=c(.4,.5,.6,.7,.8,.9), labels = c("40%","50%","60%","70%","80%","90%"))
AIRPEFig <- AIRPEFig + theme(
        text=element_text(family="sans",size=12),
        strip.background =element_rect(fill="white"),
        panel.spacing = unit(1, "lines"))
        
AIRPEFig <- AIRPEFig + labs(title="Predicted Memory Accuracy by Choice Context",
        x="Absolute Value of Prediction Error",
        y= "Estimated Marginal Means for \n% Correct Memory Responses",
        linetype="Asymmetry\nIndex",
        color = "Asymmetry\nIndex")
AIRPEFig

#2-way interaction plot (the significant effect in this model)
plot_model(lmAIRPEMaxPlot,type = "pred",
           terms = c("PositiveRPEFact", "AsymmIdx [-.8,0,.8]"), 
           colors = c("#bdd7e7","#6baed6","#2171b5"))+ 
    labs(title="Predicted Memory Accuracy by Choice Context",
        x="Absolute Value of Prediction Error",
        y= "Estimated Marginal Means for \n% Correct Memory Responses",
        linetype="Asymmetry\nIndex",
        color = "Asymmetry\nIndex")

```


# Good fit 
## subset all the dataframes to include only good fits
```{r subsetGoodFits}

#find the subs who have a correlation greater than 0 in all models, and with nloglik less than 0
corrsFull$GoodFit <- ifelse(corrsFull$correlationFull>0 & corrsFull$nloglik<0, TRUE,FALSE)

corrsFull_GF <- subset(corrsFull, corrsFull$GoodFit ==TRUE)

#mean/SD of correlation with all participants in the sample
mean(corrsFull_GF$correlationFull)
sd(corrsFull_GF$correlationFull)

#get list of subs with a good fit
GoodFit <- subset(corrsFull$subj, corrsFull$GoodFit == TRUE)

#subset AI dataset
AIDF_Full_GF <- subset(AIDF_Full, AIDF_Full$subj %in% GoodFit)

#subset Mem dataset
MemDF2_Full_GF <- subset(MemDF2_Full, MemDF2_Full$subj %in% GoodFit)

#rescaling variables because we removed some subs
MemDF2_Full_GF$AIScale <- scale(MemDF2_Full_GF$AsymmIdx)

MemDF2_Full_GF$TrialScale <- scale(MemDF2_Full_GF$trialMemory)

MemDF2_Full_GF$AbsRPEScale <- scale(MemDF2_Full_GF$AbsRPE)

MemDF2_Full_GF$FAScale <- scale(MemDF2_Full_GF$FARate)

```

## parameter distributions
```{r goodFitparams}

#plot parameter distributions

ggplot(AIDF_Full_GF, aes(x=alphapos)) + geom_histogram(fill = c("#6baed6"))+  xlab(expression(paste(alpha,"+")))

ggplot(AIDF_Full_GF, aes(x=alphaneg)) + geom_histogram(fill = c("#6baed6")) +  xlab(expression(paste(alpha,"-")))

ggplot(AIDF_Full_GF, aes(x=AI)) + geom_histogram(fill = c("#6baed6")) +xlab("Asymmetry Index (AI)")

mean(AIDF_Full_GF$AI)
sd(AIDF_Full_GF$AI)

```


## glmer
```{r glmerModelEstimatesgforig}

#run glmer model with good fit subjects

lmAIRPEMaxGF <- glmer(hit ~ AIScale*AbsRPEScale*PositiveRPEC+TrialScale+FAScale+ (1+AbsRPEScale +PositiveRPEC+TrialScale || subj), family=binomial, data=MemDF2_Full_GF,control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1e6)))

tab_model(lmAIRPEMaxGF, show.stat = TRUE, 
                           show.df = TRUE, string.stat = "z",
                           col.order = c("stat","p","est", "ci")) 


#Plot odds ratios
plotlabels <- c("Asymmetry Index (AI)", "PE Magnitude","PE Valence","Memory Trial Number","False Alarm Rate", "AI:PE Magnitude","AI:PE Valence","PE Magnitude:PE Valence","AI:PEMagnitude:PEValence")
names(plotlabels) <- c("AIScale", "AbsRPEScale","PositiveRPEC","TrialScale","FAScale","AIScale:AbsRPEScale","AIScale:PositiveRPEC","AbsRPEScale:PositiveRPEC","AIScale:AbsRPEScale:PositiveRPEC")
threewayintxplotPaper <- plot_model(lmAIRPEMaxGF, colors = "bw", show.values = TRUE, 
                                    value.offset = .4,  order.terms=c(4,5,1,2,3,6,7,8,9),
                                    title = "Fixed Effects",axis.labels = plotlabels, 
                                    vline.color = "grey")
threewayintxplotPaper + ylab("Odds of Correct Memory Response")+ theme(
        text=element_text(family="sans",size=12))  + scale_y_log10(limits = c(.5,2))
#saved manually - 600x400 pixels

#running a new model with non-centered variables to easily plot the 3-way intx

MemDF2_Full_GF$PositiveRPEFact <- factor(MemDF2_Full_GF$PositiveRPE, levels = c(0,1), labels = c("Negative Prediction Error","Positive Prediction Error"))


#run model 

lmAIRPEMaxGFPlot <- glmer(hit ~ AsymmIdx*AbsRPE*PositiveRPEFact+TrialScale+FARate+ (1+AbsRPEScale +PositiveRPEC+TrialScale || subj), family=binomial, data=MemDF2_Full_GF,control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=1e6)))

#make 3-way interaction plot
AIRPEFig <- plot_model(lmAIRPEMaxGFPlot,
           type = "pred", 
           terms = c("AbsRPE [all]", "AsymmIdx [-.8,0,.8]", "PositiveRPEFact"),
           colors = c("#bdd7e7","#6baed6","#2171b5")) +
    scale_y_continuous(breaks=c(.4,.5,.6,.7,.8,.9), labels = c("40%","50%","60%","70%","80%","90%"))
AIRPEFig <- AIRPEFig + theme(
        text=element_text(family="sans",size=12),
        strip.background =element_rect(fill="white"),
        panel.spacing = unit(1, "lines"))
        
AIRPEFig <- AIRPEFig + labs(title="Predicted Memory Accuracy by Choice Context",
        x="Absolute Value of Prediction Error",
        y= "Estimated Marginal Means for \n% Correct Memory Responses",
        linetype="Asymmetry\nIndex",
        color = "Asymmetry\nIndex")
AIRPEFig


```




